---
title: "Global Distribution of Aspergillus"
author: "Maryam Ijadpanahsaravi"
date: "`r Sys.Date()`"
output: html_document
---

### **📌 Introduction**
This document visualizes the **global distribution of Aspergillus species** using GBIF data.  
The dataset is processed, cleaned, and mapped to country-level data using **ISO3 country codes**.  

---

## ** Load Required Libraries**
```{r}
# Install missing packages if needed (uncomment the line below)
# install.packages(c("rworldmap", "classInt", "RColorBrewer", "dplyr", "countrycode", "data.table"))

# Load required libraries
library(data.table)       # For reading large CSV files
library(rworldmap)        # World map data
library(classInt)         # Classification of numeric data
library(RColorBrewer)     # Color palettes
library(dplyr)            # Data manipulation
library(countrycode)      # Country code conversion
```
## ** Load & Clean GBIF Data**
```{r}
# Load the dataset
gbif_data <- fread("Apergillus_data.csv", stringsAsFactors = FALSE, data.table = FALSE)

# Clean latitude & longitude data
gbif_data_cleaned <- gbif_data %>%
  mutate(
    decimalLatitude = as.numeric(str_replace_all(decimalLatitude, "[^0-9\\.-]", "")),  
    decimalLongitude = as.numeric(str_replace_all(decimalLongitude, "[^0-9\\.-]", ""))  
  ) %>%
  filter(
    !is.na(decimalLatitude) & !is.na(decimalLongitude),
    decimalLatitude >= -90 & decimalLatitude <= 90,
    decimalLongitude >= -180 & decimalLongitude <= 180
  )

# Convert to a standard data frame if needed
gbif_data_cleaned <- as.data.frame(gbif_data_cleaned)

```

## **Convert ISO2 Country Codes to ISO3**
```{r}
# Convert countryCode (ISO2) to ISO3
gbif_data_cleaned <- gbif_data_cleaned %>%
  mutate(iso3 = countrycode(countryCode, "iso2c", "iso3c")) %>%
  filter(!is.na(iso3))  # Remove rows where conversion failed
```
## **Aggregate Species Occurrences Per Country**
```{r}
# Count unique species per country
species_count <- gbif_data_cleaned %>%
  group_by(iso3) %>%
  summarise(total_species = n_distinct(species))

# Merge species count with cleaned GBIF data
merged_data <- left_join(gbif_data_cleaned, species_count, by = "iso3")
head(merged_data)

```
## **Match GBIF Data to World Map**
```{r}
# Load world map data
world_map <- getMap(resolution = "coarse")

# Join GBIF data with world map
sPDF <- joinCountryData2Map(merged_data, joinCode = "ISO3", nameJoinColumn = "iso3")

```
```{r}
# Convert NA values in total_species to 0 for mapping
sPDF$total_species[is.na(sPDF$total_species)] <- 0

```
## **Define Color Palette & Classification**
```{r}
# Define color palette (RdPu)
colourPalette <- brewer.pal(5, 'RdPu')

# Define class intervals using the Jenks method
classInt <- classIntervals(sPDF$total_species, n = 5, style = "jenks")
catMethod <- classInt$brks

```
## **Generate the Global Aspergillus Hotspot Map**
```{r}
# Generate the map
mapParams <- mapCountryData(sPDF, 
                            nameColumnToPlot = "total_species", 
                            addLegend = FALSE, 
                            catMethod = catMethod, 
                            colourPalette = colourPalette)

# Add a legend
do.call(addMapLegend, c(mapParams, 
                        legendLabels = "all", 
                        legendWidth = 0.5, 
                        legendIntervals = "data"))

```
##**Check for Unmatched Countries**
```{r}
# Identify GBIF country codes that did not match rworldmap
setdiff(merged_data$iso3, world_map@data$ISO_A3)

```
##**Session Info for Reproducibility**
```{r}
sessionInfo()

```

