---
title: "Global Distribution of Aspergillus by Climate Zone"
author: "Maryam Ijadpanahsaravi"
date: "`r Sys.Date()`"
output: html_notebook
---

## **📌 Introduction**
This notebook analyzes the **global distribution of Aspergillus species across different climate zones** using **GBIF occurrence data**.

---

## **1️⃣ Load Required Libraries**
```{r}
# Install missing packages (if necessary)
# install.packages(c("tidyverse", "data.table", "ggplot2", "leaflet", "dplyr"))

# Load required libraries
library(tidyverse)  # For data manipulation and visualization
library(data.table) # Efficient data loading for large files
library(ggplot2)    # For plotting
library(leaflet)    # For interactive maps
library(dplyr)      # For data manipulation
library(stringr)    # For text processing
```
2️⃣ Load & Explore the Dataset
```{r}
# Load dataset
gbif_data <- fread("Apergillus_data.csv", stringsAsFactors = FALSE, data.table = FALSE)

# Get basic info
summary(gbif_data)  # Summary statistics
str(gbif_data)      # Structure of the dataset
glimpse(gbif_data)  # A quick overview
dim(gbif_data)      # Number of rows and columns
colnames(gbif_data) # Column names

# Check for missing values
colSums(is.na(gbif_data))

# Check unique species count (if "species" column exists)
if("species" %in% colnames(gbif_data)) {
  length(unique(gbif_data$species))
} else {
  "No 'species' column found"
}

```
3️⃣ Clean Latitude & Longitude Data
```{r}
# Convert latitude & longitude to numeric and remove non-numeric characters
gbif_data_cleaned <- gbif_data %>%
  mutate(
    decimalLatitude = as.numeric(str_replace_all(decimalLatitude, "[^0-9\\.-]", "")),
    decimalLongitude = as.numeric(str_replace_all(decimalLongitude, "[^0-9\\.-]", ""))
  ) %>%
  filter(
    !is.na(decimalLatitude) & !is.na(decimalLongitude),
    between(decimalLatitude, -90, 90),
    between(decimalLongitude, -180, 180)
  )

# Convert to data.frame if necessary
gbif_data_cleaned <- as.data.frame(gbif_data_cleaned)

# Check for remaining missing values
colSums(is.na(gbif_data_cleaned[c("decimalLatitude", "decimalLongitude")]))

```
4️⃣ Visualizing Geographic Distribution
📍 Histogram of Latitude & Longitude
```{r}
# Latitude distribution
ggplot(gbif_data_cleaned, aes(x = decimalLatitude)) +
  geom_histogram(bins = 30, fill = "navy", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of Latitude", x = "Latitude", y = "Count")

# Longitude distribution
ggplot(gbif_data_cleaned, aes(x = decimalLongitude)) +
  geom_histogram(bins = 30, fill = "red3", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of Longitude", x = "Longitude", y = "Count")

```
📍 Scatterplot of Aspergillus Occurrences
```{r}
ggplot(gbif_data_cleaned, aes(x = decimalLongitude, y = decimalLatitude)) +
  geom_point(alpha = 0.5, color = "navy") +
  theme_minimal() +
  labs(title = "Geographic Distribution of Aspergillus", x = "Longitude", y = "Latitude")

```
5️⃣ Interactive Map of Aspergillus Occurrences
```{r}
leaflet(gbif_data_cleaned) %>%
  addTiles() %>%
  addCircleMarkers(
    ~decimalLongitude, ~decimalLatitude,
    radius = 2, color = "navy", stroke = FALSE, fillOpacity = 0.7
  ) %>%
  setView(lng = mean(gbif_data_cleaned$decimalLongitude, na.rm = TRUE),
          lat = mean(gbif_data_cleaned$decimalLatitude, na.rm = TRUE),
          zoom = 2)

```
6️⃣ Classifying Climate Zones
```{r}
# Define climate zones based on latitude
gbif_data_cleaned <- gbif_data_cleaned %>%
  mutate(
    climate_zone = case_when(
      between(decimalLatitude, -23.5, 23.5)  ~ "Tropical",
      between(decimalLatitude, -35, -23.5) | between(decimalLatitude, 23.5, 35) ~ "Subtropical",
      between(decimalLatitude, -60, -35) | between(decimalLatitude, 35, 60) ~ "Temperate",
      decimalLatitude > 60 | decimalLatitude < -60  ~ "Polar",
      TRUE ~ "Unknown"
    )
  )

# Count records per climate zone
table(gbif_data_cleaned$climate_zone)

```
7️⃣ Visualizing Aspergillus Abundance in Climate Zones
```{r}
# Count occurrences per climate zone
climate_counts <- gbif_data_cleaned %>%
  group_by(climate_zone) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Bar plot
ggplot(climate_counts, aes(x = reorder(climate_zone, -count), y = count, fill = climate_zone)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Frequency of Aspergillus Occurrences in Climate Zones",
       x = "Climate Zone",
       y = "Number of Occurrences") +
  scale_fill_manual(values = c("Tropical" = "#FF9900", 
                               "Subtropical" = "#FFCC00",
                               "Temperate" = "#0099CC", 
                               "Polar" = "#003366")) +
  theme(legend.position = "none")

```

Tropical regions have the highest Aspergillus occurrences, likely due to warm, humid conditions favorable for fungal growth.
Temperate zones also show significant occurrence levels.
Subtropical & Polar regions have the least reported cases, possibly due to lower fungal survival rates or less sampling effort.